---
title: "Week 3 - Homework"
author: "STAT 420, Summer 2018, Xin Qu"
date: ''
output:
  html_document: 
    toc: yes
  pdf_document: default
urlcolor: cyan
---
***

## Exercise 1 (Using `lm` for Inference)

For this exercise we will use the `cats` dataset from the `MASS` package. You should use `?cats` to learn about the background of this dataset.

**(a)** Fit the following simple linear regression model in `R`. Use heart weight as the response and body weight as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `cat_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.
```{r}
library(MASS)
cat_model = lm(Hwt ~ Bwt, data = cats)
summary(cat_model)
```

- The null and alternative hypotheses

For $\beta_0$: $H_0:\beta_0 = 0$ vs $H_1:\beta_0\neq 0$
For $\beta_1$: $H_0:\beta_1 = 0$ vs $H_1:\beta_1\neq 0$

- The value of the test statistic

For $\beta_0$
```{r}
summary(cat_model)$coefficients[1, 3]
```

For $\beta_1$
```{r}
summary(cat_model)$coefficients[2, 3]
```


- The p-value of the test

For $\beta_0$
```{r}
summary(cat_model)$coefficients[1, 4]
```

For $\beta_1$
```{r}
summary(cat_model)$coefficients[2, 4]
```


- A statistical decision at $\alpha = 0.05$

For $\beta_0$
```{r}
summary(cat_model)$coefficients[1, 4] < 0.05
```

For $\beta_1$
```{r}
summary(cat_model)$coefficients[2, 4] < 0.05
```

**Since the p-value of $\beta_1$ is much smaller than $\alpha = 0.05$, reject the null hypothesis $H_0:\beta_1=0$. The p-value of $\beta_0$ is larger than $\alpha = 0.05$, fail to reject the null hypothesis $H_0:\beta_0=0$.**

- A conclusion in the context of the problem

**There is a significant linear relationship between heart weight and body weight of cats in the dataset.**

**(b)** Calculate a 90% confidence interval for $\beta_1$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(cat_model, parm = 'Bwt', level = 0.9)
```

**Give an interpretation of the interval in the context of the problem.**

**We are 90% confident that when the body weight increses 1 kg, the increase of the mean heart weight is in range [3.619716, 4.448409] g (inclusive).**

**(c)** Calculate a 99% confidence interval for $\beta_0$. Give an interpretation of the interval in the context of the problem.

```{r}
confint(cat_model, parm = '(Intercept)', level = 0.99)
```

**Give an interpretation of the interval in the context of the problem.**

**We are 99% confident that the mean heart weight for a 0 kg body weight is in range [-2.164125, 1.4508] (inclusive) g. But it's unlikely to have a 0 kg body weight of a cat and the heart weight is not supposed to be a negative number.**

**(d)** Use a 99% confidence interval to estimate the mean heart weight for body weights of 2.1 and 2.8 kilograms. Which of the two intervals is wider? Why?

```{r}
intervals = predict(cat_model, newdata = data.frame(Bwt = c(2.1, 2.8)), interval = 'confidence', level = 0.99)
intervals[1, 3] - intervals[1, 2]
intervals[2, 3] - intervals[2, 2]
```

**Which of the two intervals is wider? Why?**
```{r}
mean(cats$Bwt)
```

**The confidence interval for 2.1kg Bwt is wider. From the above, we could tell the mean of Bwt is 2.72kg and 2.1 is further away from the mean than 2.8 kg, which means the SE becomes larger.**

**(e)** Use a 99% prediction interval to predict the heart weight for body weights of 2.8 and 4.2 kilograms.
```{r}
predict(cat_model, newdata = data.frame(Bwt = c(2.8, 4.2)), interval = 'prediction', level = 0.99)
```

**(f)** Create a scatterplot of the data. Add the regression line, 90% confidence bands, and 90% prediction bands.

```{r}
Bwt_grid = seq(min(cats$Bwt), max(cats$Bwt), by = 0.01)
Hwt_ci_band = predict(cat_model, newdata = data.frame(Bwt = Bwt_grid), interval = 'confidence', level = 0.9)
Hwt_pi_band = predict(cat_model, newdata = data.frame(Bwt = Bwt_grid), interval = 'prediction', level = 0.9)
plot(Hwt ~ Bwt, data = cats, xlab = 'Body weight (kg)', 
     ylab = 'Heart weight (g)', main = 'Heart weight VS Body weight', 
     pch = 20, cex = 2, col = 'grey', ylim = c(min(Hwt_pi_band), max(Hwt_pi_band)))
abline(cat_model, lwd = 5, col = 'orange')
lines(Bwt_grid, Hwt_ci_band[, 'lwr'], col = 'blue', lty = 2, lwd = 3)
lines(Bwt_grid, Hwt_ci_band[, 'upr'], col = 'blue', lty = 2, lwd = 3)
lines(Bwt_grid, Hwt_pi_band[, 'lwr'], col = 'blue', lty = 3, lwd = 3)
lines(Bwt_grid, Hwt_pi_band[, 'upr'], col = 'blue', lty = 3, lwd = 3)
```

**(g)** Use a $t$ test to test:

- $H_0: \beta_1 = 4$
- $H_1: \beta_1 \neq 4$

Report the following:

- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.05$

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

- The value of the test statistic
```{r}
t = (summary(cat_model)$coefficients[2, 1]- 4) / summary(cat_model)$coefficients[2, 2]
t
#summary(cat_model)$coefficients[2, 3] * (summary(cat_model)$coefficients[2, 1] - 4) / summary(cat_model)$coefficients[2, 1]
```

- The p-value of the test
```{r}
2 * pt(t, df = nrow(cats) - 2, lower.tail = FALSE)
```

- A statistical decision at $\alpha = 0.05$

From the above, the p-value is larger than $\alpha = 0.05$, we fail to reject the null hypothesis. 
***

## Exercise 2 (More `lm` for Inference)

For this exercise we will use the `Ozone` dataset from the `mlbench` package. You should use `?Ozone` to learn about the background of this dataset. You may need to install the `mlbench` package. If you do so, do not include code to install the package in your `R` Markdown document.

For simplicity, we will re-perform the data cleaning done in the previous homework.

```{r}
data(Ozone, package = "mlbench")
Ozone = Ozone[, c(4, 6, 7, 8)]
colnames(Ozone) = c("ozone", "wind", "humidity", "temp")
Ozone = Ozone[complete.cases(Ozone), ]
```

**(a)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and wind speed as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_wind_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.

```{r}
ozone_wind_model = lm(ozone ~ wind, data = Ozone)
summary(ozone_wind_model)
```


- The null and alternative hypotheses

$H_0: \beta_0 = 0$ vs $H_1: \beta_0 \neq 0$

$H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$

- The value of the test statistic

For $\beta_0$
```{r}
summary(ozone_wind_model)$coefficient[1, 3]
```

For $\beta_1$
```{r}
summary(ozone_wind_model)$coefficient[2, 3]
```
- The p-value of the test

For $\beta_0$
```{r}
summary(ozone_wind_model)$coefficient[1, 4]
```

For $\beta_1$
```{r}
summary(ozone_wind_model)$coefficient[2, 4]
```

- A statistical decision at $\alpha = 0.01$

Since the p-value of $\beta_1$ is larger than $\alpha$, we fail to reject the null hypothesis$H_0: \beta_1 = 0$.

- A conclusion in the context of the problem

There is not a significant linear relationship between the ozone measurement and wind speed.

**(b)** Fit the following simple linear regression model in `R`. Use the ozone measurement as the response and temperature as the predictor. 

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Store the results in a variable called `ozone_temp_model`. Use a $t$ test to test the significance of the regression. Report the following:

- The null and alternative hypotheses
- The value of the test statistic
- The p-value of the test
- A statistical decision at $\alpha = 0.01$
- A conclusion in the context of the problem

When reporting these, you should explicitly state them in your document, not assume that a reader will find and interpret them from a large block of `R` output.
```{r}
ozone_temp_model = lm(ozone ~ temp, data = Ozone)
summary(ozone_temp_model)
```

- The null and alternative hypotheses

$H_0: \beta_0 = 0$ vs $H_1: \beta_0 \neq 0$

$H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$

- The value of the test statistic

For $\beta_0$
```{r}
summary(ozone_temp_model)$coefficient[1, 3]
```

For $\beta_1$
```{r}
summary(ozone_temp_model)$coefficient[2, 3]
```
- The p-value of the test

For $\beta_0$
```{r}
summary(ozone_temp_model)$coefficient[1, 4]
```

For $\beta_1$
```{r}
summary(ozone_temp_model)$coefficient[2, 4]
```

- A statistical decision at $\alpha = 0.01$

Since the p-value of $\beta_1$ is much smaller than $\alpha$, we reject the null hypothesis$H_0: \beta_1 = 0$.

- A conclusion in the context of the problem

There is a significant linear relationship between the ozone measurement and temperature.

***

## Exercise 3 (Simulating Sampling Distributions)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = -5$
- $\beta_1 = 3.25$
- $\sigma^2 = 16$

We will use samples of size $n = 50$.

**(a)** Simulate this model $2000$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_0$ and $\hat{\beta}_1$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 19890310
set.seed(birthday)
n = 50
x = seq(0, 10, length = n)
```

```{r}
num_times = 2000
beta_0 = -5
beta_1 = 3.25
sigma = sqrt(16)
beta_0_hats = rep(0, num_times)
beta_1_hats = rep(0, num_times)
for (i in 1: num_times) {
  epsilon = rnorm(n, mean = 0, sd = sigma)
  y = beta_0 + beta_1 * x + epsilon
  
  sim_model = lm(y ~ x)
  beta_0_hats[i] = coef(sim_model)[1]
  beta_1_hats[i] = coef(sim_model)[2]
}
```

**(b)** Create a table that summarizes the results of the simulations. The table should have two columns, one for $\hat{\beta}_0$ and one for $\hat{\beta}_1$. The table should have four rows:

- A row for the true expected value given the known values of $x$
- A row for the mean of the simulated values
- A row for the true standard deviation given the known values of $x$
- A row for the standard deviation of the simulated values

```{r}
Sxx = sum((x - mean(x)) ^ 2)
result = data.frame('Variables' = c('True expected value', 'Mean of simulated values', 'True sd', 'Sd of simulated values'),
                    'Beta_0' = c(beta_0, mean(beta_0_hats), sqrt(sigma ^ 2 * (1 / n + mean(x) ^ 2 / Sxx)), sd(beta_0_hats)), 
                    'Beta_1' = c(beta_1, mean(beta_1_hats), sqrt(sigma ^ 2 / Sxx), sd(beta_1_hats)))
library(knitr)
kable(result, digits = 4)
```

**(c)** Plot two histograms side-by-side:

- A histogram of your simulated values for $\hat{\beta}_0$. Add the normal curve for the true sampling distribution of $\hat{\beta}_0$.
- A histogram of your simulated values for $\hat{\beta}_1$. Add the normal curve for the true sampling distribution of $\hat{\beta}_1$.

```{r}
par(mfrow=c(1,2))
hist(beta_0_hats, probability = TRUE, breaks = 20, xlab = expression(hat(beta)[0]),
     main = '', border = 'blue')
curve(dnorm(x, mean = beta_0, sd = sqrt(sigma ^ 2 * (1 / n + mean(x) ^ 2 / Sxx))), col = 'orange', 
      add = TRUE, lwd = 3)
hist(beta_1_hats, probability = TRUE, breaks = 20, xlab = expression(hat(beta)[1]),
     main = '', border = 'darkgrey')
curve(dnorm(x, mean = beta_1, sd = sqrt(sigma ^ 2 / Sxx)), col = 'orange', 
      add = TRUE, lwd = 3)
```


***

## Exercise 4 (Simulating Confidence Intervals)

For this exercise we will simulate data from the following model:

\[
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i
\]

Where $\epsilon_i \sim N(0, \sigma^2).$ Also, the parameters are known to be:

- $\beta_0 = 5$
- $\beta_1 = 2$
- $\sigma^2 = 9$

We will use samples of size $n = 25$.

Our goal here is to use simulation to verify that the confidence intervals really do have their stated confidence level. Do **not** use the `confint()` function for this entire exercise.

**(a)** Simulate this model $2500$ times. Each time use `lm()` to fit a simple linear regression model, then store the value of $\hat{\beta}_1$ and $s_e$. Set a seed using **your** birthday before performing the simulation. Note, we are simulating the $x$ values once, and then they remain fixed for the remainder of the exercise.

```{r}
birthday = 19893010
set.seed(birthday)
n = 25
x_4 = seq(0, 2.5, length = n)
```
```{r}
beta_0_4 = 5
beta_1_4 = 2
sigma_4 = sqrt(9)
Sxx_4 = sum((x_4 - mean(x_4)) ^ 2)
beta_1_hats_4 = rep(0, 2500)
beta_1_hats_se = rep(0, 2500)

for (i in 1: 2500) {
  epsilon_4 = rnorm(n, mean = 0, sd = sigma_4)
  y_4 = beta_0_4 + beta_1_4 * x_4 + epsilon_4
  sim_model_4 = lm(y_4 ~ x_4)
  beta_1_hats_4[i] = coef(sim_model_4)[2]
  beta_1_hats_se[i] = summary(sim_model_4)$sigma / sqrt(Sxx_4)
}
```

**(b)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 95% confidence interval. Store the lower limits in a vector `lower_95` and the upper limits in a vector `upper_95`. Some hints:

- You will need to use `qt()` to calculate the critical value, which will be the same for each interval.
- Remember that `x` is fixed, so $S_{xx}$ will be the same for each interval.
- You could, but do not need to write a `for` loop. Remember vectorized operations.
```{r}
alpha = 1 - 0.95
crit_value = qt(1 - alpha / 2, df = 25 - 2)
lower_95 = beta_1_hats_4 - crit_value * beta_1_hats_se
upper_95 = beta_1_hats_4 + crit_value * beta_1_hats_se
```

**(c)** What proportion of these intervals contains the true value of $\beta_1$?
```{r}
#sum(as.numeric(lower_95 <= beta_1_4 & beta_1_4 <= upper_95)) / 2500
mean(lower_95 <= beta_1_4 & beta_1_4 <= upper_95)
```

**(d)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.05$?
```{r}
1 - mean(lower_95 <= 0 & 0 <= upper_95)
```

**(e)** For each of the $\hat{\beta}_1$ that you simulated, calculate a 99% confidence interval. Store the lower limits in a vector `lower_99` and the upper limits in a vector `upper_99`.

```{r}
alpha_e = 1 - 0.99
crit_value_e = qt(1 - alpha_e / 2, df = 25 - 2)
lower_99 = beta_1_hats_4 - crit_value_e * beta_1_hats_se
upper_99 = beta_1_hats_4 + crit_value_e * beta_1_hats_se
```


**(f)** What proportion of these intervals contains the true value of $\beta_1$?
```{r}
mean(lower_99 <= beta_1_4 & beta_1_4 <= upper_99)
```

**(g)** Based on these intervals, what proportion of the simulations would reject the test $H_0: \beta_1 = 0$ vs $H_1: \beta_1 \neq 0$ at $\alpha = 0.01$?

```{r}
1 - mean(lower_99 <= 0 & 0 <= upper_99)
```

***

## Exercise 5 (Prediction Intervals "without" `predict`)

Write a function named `calc_pred_int` that performs calculates prediction intervals:

$$
\hat{y}(x) \pm t_{\alpha/2, n - 2} \cdot s_e\sqrt{1 + \frac{1}{n}+\frac{(x-\bar{x})^2}{S_{xx}}}.
$$

for the linear model

$$
Y_i = \beta_0 + \beta_1 x_i + \epsilon_i.
$$

**(a)** Write this function. You may use the `predict()` function, but you may **not** supply a value for the `level` argument of `predict()`. (You can certainly use `predict()` any way you would like in order to check your work.)

The function should take three inputs:

- `model`, a model object that is the result of fitting the SLR model with `lm()`
- `newdata`, a data frame with a single observation (row)
    - This data frame will need to have a variable (column) with the same name as the data used to fit `model`.
- `level`, the level (0.90, 0.95, etc) for the interval with a default value of `0.95`

The function should return a named vector with three elements:

- `estimate`, the midpoint of the interval
- `lower`, the lower bound of the interval
- `upper`, the upper bound of the interval
```{r}
calc_pred_int = function(model, newdata, level = 0.95) {
  estimate = predict(model, newdata = data.frame(newdata))
  Se = summary(model)$sigma
  alpha_5 = 1 - level
  n = length(summary(model)$residuals)
  crit_value_5 = qt(1 - alpha_5 / 2, df = n - 2)
  x_5 = model$model[, 2]
  Sxx_5 = sum((x_5 - mean(x_5)) ^ 2)
  lower = estimate - crit_value_5 * Se * sqrt(1 / n + (newdata - mean(x_5)) ^ 2 / Sxx_5 + 1)
  upper = estimate + crit_value_5 * Se * sqrt(1 / n + (newdata - mean(x_5)) ^ 2 / Sxx_5 + 1)
  v = c(estimate, lower, upper)
  names(v) = c('Estimate', 'Lower', 'Upper')
  return(v)
}
```

**(b)** After writing the function, run this code:

```{r}
newcat_1 = data.frame(Bwt = 4.0)
calc_pred_int(cat_model, newcat_1)
```


```{r, eval=FALSE}
predict(cat_model, newdata = data.frame(Bwt = 4.0), interval = 'prediction', level = 0.95)
```

**(c)** After writing the function, run this code:

```{r}
newcat_2 = data.frame(Bwt = 3.3)
calc_pred_int(cat_model, newcat_2, level = 0.99)
```

```{r, eval=FALSE}
predict(cat_model, newdata = data.frame(Bwt = 3.3), interval = 'prediction', level = 0.99)
```


